---
title: "choose your own project"
author: "John Kingsley"
date: "2019/12/24"
output: html_document
---
INTRODUCTION
Video games are widely played all over the world and the industries in this area of business have tremendously grown in the past decades. Video games are electronic designed algorithm that are incorporated into a computing device which includes computers (either desktop or laptop), mobile phone or a gaming console. Video games are subdivided into mobile games and computers, which solely dependent on the platforms. In the early 80s and late 70s, there were two major markets for video games, the home and the arcade markets. The arcade market generated approximately 8 billion USD more than pop music as at 1982 drawing huge attention from investors. The video games have then grown tremendously and the games enjoyed all over the world by all class of people irrespective of their age. More so, some video games have come and gone while others still in the industry are developing modern games with the needs of the people and with the growing technologies
VIDEO GAMES INDUSTRY PROJECTION
In 2016, The market research conducted by the Newzoo firm showed that it predicts the global game market to grow up to $99.6 billion which was about 8.4 percent when compared to that of the previous year. In this year, the firm also predicted that mobile games will experience more sales for the first time over the console games industry with 21.3 percent growth when compared to previous year.
TRENDS IN THE INDUSTRY OVER THE YEARS
1. Sales of games are dominant in the North America region
2.puzzled games has obviously declined in the game popularity while Action and adventure games have experienced positive growth.

PROJECT OVERVIEW
The main aim of the project is to show relationshiop in sales
Other aims includes
a. Sales at different regions of the world
b. Popualar genre at global sales
c. Popular Publishers at global sales
d The year with the highest number of sales.
a. to see if there are relationships between sales of the regions
b. to see if there are relationship between sales and genres in the different regions of the world
This we can achieve through data visualization after that we can further do the following data analysis, to note observable changes in comparing the genre of a game and the platform which they are released in, to ascertain if there are any relationship between sales in the regions, genres and sales.
KEYS TO NOTE: Name: Name of the video game

Platform: Platform on which the game was released or is playable

Year: Year in which the game was released

Genre: Genre the game belongs to

Publisher: Name of the publisher who created the game

NA_Sales: Sales in North America

EU_Sales: Sales in Europe

JP_Sales: Sales in Japan

Other_Sales: Sales in other countries

Global_Sales: Global Sales
DATA
The data for the project was obtained from (https://www.kaggle.com/rush4ratio/video-game-sales-with-ratings/data#Video_Games_Sales_as_at_22_Dec_2016.csv) which was already in it tidy state and available for machine learning analsysis. The data provided insight on the trends of video game industrie for over 30 years in the different regions of the world and also at global perspective.

EXECUIVE SUMMARY
The analysis offered me great insight into the details of the video game industry which are widely played and enjoyed by all regions of the world regardless of age. The analysis revealed that there is a relationship between regional sales and globall. That the increase or decrease in regional sales will proportional influence global sales. And that North America regions are the highest consumers of video games followed by European Union. Similarly, Shooter, Sport and Action genres have maintained top spot in the video game industry over the years. And that DS, PS, XBOX and Wii are in the top spots of the video game industry.

METHDOLOGY AND ANALYSIS
#load packages
```{r}
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")
library(viridis)
library(wordcloud)
library(RColorBrewer)
library(magrittr)
library(lubridate)
library(RPostgreSQL)
library(plotly)
library(jsonlite)
library(htmltools)
library(glmnet)
library(epitools)
library(lme4)
library(sjPlot)
library(pscl)
```

#loading of dataset
```{r}
VIDEOGS <- read.csv("D:/john/Videogs_2016.csv")
VIDEOGS
```

#data inspection
```{r}
head(VIDEOGS)
glimpse(VIDEOGS)
summary(VIDEOGS)
names(VIDEOGS)
dim(VIDEOGS)
```

#class of some variables
```{r}
class(VIDEOGS$Name)
class(VIDEOGS$Platform)
class(VIDEOGS$Publisher)
class(VIDEOGS$Year_of_Release)
class(VIDEOGS$Genre)
```

#load package
```{r}
library(psych)
describe(VIDEOGS)
```

#Sales trend around the regions of the world according to platforms, genres, names and publishers.
#Sales in North America
```{r}
boxplot(VIDEOGS$NA_Sales, main="Sales in North America", xlab="Sales", ylab="frequency", vertical = TRUE)
summary(VIDEOGS$NA_Sales)
```

#sales in Europe
```{r}
boxplot(VIDEOGS$EU_Sales, main="Sales in Europe", xlab="Sales", ylab="frequency", vertical = TRUE)
summary(VIDEOGS$EU_Sales)
```


#sales in Japan
```{r}
boxplot(VIDEOGS$JP_Sales, main="Sales in Japan", xlab="Sales", ylab="frequency", vertical = TRUE)
summary(VIDEOGS$JP_Sales)
```

#other sales
```{r}
boxplot(VIDEOGS$Other_Sales, main="Sales in Japan", xlab="Sales", ylab="frequency", vertical = TRUE)
summary(VIDEOGS$JP_Sales)
```

#Global sales
```{r}
boxplot(VIDEOGS$Global_Sales, main="Sales in Japan", xlab="Sales", ylab="frequency", vertical = TRUE)
summary(VIDEOGS$Global_Sales)
```

#Popular Genre by sales at global level
```{r}
VIDEOGS%>%
  group_by(Genre) %>%
  summarise(Count = n()) %>%
  plot_ly(x = ~Genre,
          y = ~Count,
          type = "bar", col="black")
```

#From the bar chart above, Action Genre games yielded the highest sales globally.
##DC and PS were the platforms that benefitted with the highest sales at the global level. Although X360 and WiiU also strived at the global the global sales.
```{r}
VIDEOGS%>%
  group_by(Platform) %>%
  summarise(Count = n()) %>%
  plot_ly(x = ~Platform,
          y = ~Count,
          type = "bar")
```

#Sales by publishers
#The bar chart revealed that Electronic Arts was the publisher with the highest sales.
```{r}
VIDEOGS%>%
  group_by(Publisher) %>%
  summarise(Count = n()) %>%
  plot_ly(x = ~Publisher,
          y = ~Count,
          type = "bar")
```

#top sales by Platforms by region
#The top platforms by sales in different region were DS, PS, PS2, PS3, WiiU and X360
```{r}
VIDEOGS %>%
  gather("Region", "Value", c("NA_Sales", "EU_Sales", "JP_Sales", "Other_Sales")) %>%
  group_by(Region, Platform) %>%
  summarize(Sales = sum(Value)) %>%
  top_n(n = 3) %>%
  ggplot(aes(x = Region, y = Sales, group = Region, fill = Platform)) +
  geom_col(position = "stack") +
  scale_fill_viridis(discrete = TRUE) +
  labs(title = "Top Genre by Sales per Region")
```

#Top sold games in the regions of the world
```{r}
VIDEOGS %>%
  gather("Region", "Value", c("NA_Sales", "EU_Sales", "JP_Sales", "Other_Sales")) %>%
  group_by(Region, Name) %>%
  summarize(Sales = sum(Value)) %>%
  top_n(n = 3) %>%
  ggplot(aes(x = Region, y = Sales, group = Region, fill = Name)) +
  geom_col(position = "stack") +
  scale_fill_viridis(discrete = TRUE) +
  labs(title = "Top Genre by Sales per Region")
```

#Top sold genres according to regions
#Action, Role-playing, Shooter and Sport
```{r}
VIDEOGS %>%
  gather("Region", "Value", c("NA_Sales", "EU_Sales", "JP_Sales", "Other_Sales")) %>%
  group_by(Region, Genre) %>%
  summarize(Sales = sum(Value)) %>%
  top_n(n = 3) %>%
  ggplot(aes(x = Region, y = Sales, group = Region, fill = Genre)) +
  geom_col(position = "stack") +
  scale_fill_viridis(discrete = TRUE) +
  labs(title = "Top Genre by Sales per Region")
```

#total sales per year at global level
```{r}
tot_year <- aggregate(VIDEOGS$Global_Sales, by=list(Year=VIDEOGS$Year), sum)
plot(tot_year)
data_frame(tot_year)
```

#Global Sales per genre
```{r}
Glb_sales <- aggregate(VIDEOGS$Global_Sales, by=list(Genre=VIDEOGS$Genre), sum)
data.table(Glb_sales)
plot(Glb_sales)
```

#number of games released per
#the output revealed that the highest number of games was produced in 2008 with a total of 1427 games which was followed by the year 2007 with 1426.
```{r}
VIDEOGS %>%
  group_by(Year_of_Release) %>%
  summarize(Number_of_games_each_year = n())
```

#PLOT
```{r}
VIDEOGS %>%
  group_by(Year_of_Release) %>%
  summarize(Number_of_games_each_year = n()) %>%
  ggplot(aes(x = Year_of_Release, y = Number_of_games_each_year)) +
  geom_col(fill = "red") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Games released per Year", x = "Year", y = "Sales (units)")
```


#a simple correlation matrix was carried out between sales
```{r}
library(corrplot)
VIDEOGS[, c("NA_Sales", "EU_Sales", "JP_Sales", "Other_Sales", "Global_Sales")] %>%
  cor(method = "pearson") %>%
  corrplot::corrplot(addCoef.col = "white", type="upper")
```

#We considered the relationship between Sales in other regions against the global sales.
```{r}
par(mfrow=c(1,4))
with(VIDEOGS, plot(NA_Sales, Global_Sales))
with(VIDEOGS, plot(EU_Sales, Global_Sales))
with(VIDEOGS, plot(JP_Sales, Global_Sales))
with(VIDEOGS, plot(Other_Sales, Global_Sales))
```

#The result revealed that there is linear relationship between regional sales and global sales.
#we further considered linear models between sales in all the regions and genres in the study since action Genres have dorminated sales.
#North America
```{r}

fit <- lm( NA_Sales ~ Genre, VIDEOGS)
summary(fit)
plot(fit)
```

#linear relationship between sales in Europe and Genre
```{r}
fit <- lm( EU_Sales ~ Genre, VIDEOGS)
summary(fit)
plot(fit)
```

#linear relationship between sales in Japan and Genre
```{r}
fit <- lm( JP_Sales ~ Genre, VIDEOGS)
summary(fit)
plot(fit)
```

#linear relationship between other sales and Genre
```{r}
fit <- lm(Other_Sales ~ Genre, VIDEOGS)
summary(fit)
plot(fit)
```
#linear relationship between global sales and Genre
```{r}
fit <- lm(Global_Sales ~ Genre, VIDEOGS)
summary(fit)
plot(fit)
```
#with reference to the fitted plots, it is obvious that genres has no influence on the sales of games in all the regions.
#we then looked at the relationship between sales and Genre and Rating
#The result revealed a strong relationship between sales and Genre. 
```{r}
VIDEOGS %>%
  ggplot(aes(x = Rating, y = Genre, col = Genre)) +
  geom_jitter(alpha = 0.6, pch = 25) +
  theme(legend.position = "none") +
  scale_color_viridis(discrete = TRUE)
```

